The ChEMBL database as Linked Open Data

### 2012-04-11 article.tex
Drafted the LOD section

\begin{footnotesize}
\begin{verbatim}
target:t1
 a chembl:Target ;
 rdfs:subClassOf pro:PR_000000001 ;
 rdfs:label "Glucoamylase" , 
  "Maltase-glucoamylase, intestinal" ;
 dc:identifier "uniprot:O43451" ,
   "3.2.1.3" ;
 dc:title "Maltase-glucoamylase" ;
 chembl:classL1 "Enzyme" ;
 chembl:hasDescription
  "Maltase-glucoamylase, intestinal" ;
 chembl:hasKeyword "Glycosidase" , 
  "Membrane" , "Sulfation" ;
 chembl:hasTaxonomy
  <http://bio2rdf.org/taxonomy:9606> ;
 chembl:organism "Homo sapiens" .
\end{verbatim}
\end{footnotesize}

@@

To make the data a proper citizen of the semantic web, we link out to various resources.
The web resources linked to as well as the predicate used for that link is given in
Figure~\ref{2}.
Links out are made to ChemSpider and RDF OpenMolecules for compounds, Bio2RDF for
protein targets, and CrossRef and Bio2RDF for DOIs and PubMed identifiers respectively.

@@

Peter/Egon: write up links to Bio2RDF for protein, pubmed, ...

@@

\begin{tiny}
\begin{verbatim}
<http://rdf.chemspider.com/370>
 skos:exactMatch
 <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
\end{verbatim}
\end{tiny}

@@

@@

>>>>

\begin{footnotesize}
\begin{verbatim}
target:t1
 a chembl:Target ;
 rdfs:subClassOf pro:PR_000000001 ;
 rdfs:label "Glucoamylase" , 
  "Maltase-glucoamylase, intestinal" ;
 dc:identifier "uniprot:O43451" ,
   "3.2.1.3" ;
 dc:title "Maltase-glucoamylase" ;
 chembl:classL1 "Enzyme" ;
 chembl:hasDescription
  "Maltase-glucoamylase, intestinal" ;
 chembl:hasKeyword "Glycosidase" , 
  "Membrane" , "Sulfation" ;
 chembl:organism "Homo sapiens" .
\end{verbatim}
\end{footnotesize}

@@

To make the data a proper citizen of the semantic web, we link out to various resources,
which is visualized in Figure~\ref{2}. Triples for compounds link out to ChemSpider 
using the new complementary index and RDF OpenMolecules using InChI values, to Bio2RDF
for protein targets using the uniprot identifier, and CrossRef and Bio2RDF for literature
references using the DOIs and PubMed identifiers, respectively.

@@

For protein links out are based on the FOOBAR for the species, the EC code for proteins,
as well as UniProt identifiers.

FIXME: should we link out to the new UniProt RDF?

\begin{tiny}
\begin{verbatim}
target:t101191
 chembl:hasTaxonomy <http://bio2rdf.org/taxonomy:9606> ;
 owl:sameAs <http://bio2rdf.org/ec:2.7.11.1> ;
 owl:sameAs <http://bio2rdf.org/uniprot:Q8IYT8> .
\end{verbatim}
\end{tiny}

For papers with PubMed identifiers we also link out to Bio2RDF:

\begin{tiny}
\begin{verbatim}
res:r23 rdfs:seeAlso <http://bio2rdf.org/pubmed:15149661> .
\end{verbatim}
\end{tiny}

@@

\begin{tiny}
\begin{verbatim}
<http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846>
 skos:exactMatch <http://rdf.chemspider.com/370> .
\end{verbatim}
\end{tiny}

@@

The InChI is a unique identifier for (small) organic molecules, and has been previously used
to define unique IRIs for molecules~\cite{???}. While IRIs are theoretically unlimited in length,
in practice web browsers and servers do limit the length of IRIs. Virtuoso is, unfortunately,
a system which supports only IRIs of up to a certain length. Therefore, only for smaller molecules
InChI-based links are created. Almost 1.3 million links are created, looking like:

\begin{tiny}
\begin{verbatim}
mol:m62687 owl:equivalentClass
 <http://rdf.openmolecules.net/?InChI=1S/CH4/h1H4> .
\end{verbatim}
\end{tiny}

Notice here the use of owl:equivalentClass to match the formalism in CHEMINF that defines molecules
are classes, rather than instances.

@@

Besides the PubMed identifiers used to link from literature references to Bio2RDF (see earlier),
ChEMBL uses DOIs alternatively, which we use to link out to the RDF provided by CrossRef~\citep{CrossRefRDF}:

\begin{tiny}
\begin{verbatim}
res:r2032 owl:sameAs <http://dx.doi.org/10.1016/0960-894X(96)00111-4> .
\end{verbatim}
\end{tiny}
